The benefits of an extended feature space in a machine learning bot for the game Schnapsen

### 2018-01-31 Text/paper.tex
Updated Experimental setup (still WIP)

\subsection{Intelligent Systems Schnapsen Framework}
An API in Python has been provided, together with a framework for the Schnapsen game. Furthermore, we  have two scripts: one where bots can play against each other in a single game where all moves in the game can be seen, and a script where bots can play a tournament against each other - with a specifiable number of the number of games to be played - and where the number of won games per robot can be seen. There are already multiple bots already present in the framework to play against:
\begin{description}
\item [Minimax / Alphabeta:] This bot only works in a perfect-information game (phase 2 of Schnapsen). It recursively looks through all the states, and decides what the best move to play will be, provided that the opponent plays optimally, and will try to minimize the outcome of the game for the player.
\item [Bully:] Plays by a defined set of rules. If it has a trump card, it will play that. Otherwise, it will play a card of the same suit player 1 played. If it cannot do either of those things, it will play the highest ranking card of any suit.
\item [Rand:] Chooses a random move out of all of the legal moves, and plays that.
\item [Rdeep:] Functions quite like a Monte-Carlo sampling bot. It looks through multiple states, and averages the heuristics of a state, thereby ranking the moves. It then picks one of the highest ranking moves to play.
\item [KBbot:] Logical reasoning bot. The strategy can be defined in propositional logic. The current strategy of the bot is to always play cheap cards (i.e kings, queens, jacks) first.
\end{description}

\subsection{Machine Learning Bots}
For this paper, we will use machine learning bots with different set features to determine what the best features are.

@@

WORK IN PROGRESSSS

To answer our research question, multiple Schnapsen tournaments were held. Each learning bot played multiple 120 game tournaments against each individual non-learning bot. Each learning bot played a total of 9 tournaments with 3 different models against each static bot.
\subsection{The bots}
The non- learning opponents were Rand, Bully and Rdeep. The learning components we created were:
\begin{description}
\item [ml\_minimal:] asdf
\item [ml\_stripped:]  asdf
\item [ml:] This ml bot is provided by the framework and consists of the  \textit{default feature space} 
\item [ml\_enriched:]  asdf
\item [ml\_advanced:] asdf
\end{description}

The research procedure was as follows:
\begin{itemize}
\item Training
\item 120 game tournament
\item 120 game tournament
\item 120 game tournament
\end{itemize}

This procedure was repeated twice to reach the total of 9 tournaments respectively. The machine-learning bots did not compete against eachother.
Both ML, and MLAdvanced were trained on the same bot. First, bully, then rand, then rdeep, then kbbot, etc.

>>>>

\subsection{Intelligent Systems Schnapsen Framework}
An API in Python has been provided, together with a framework for the Schnapsen game. Furthermore, we  have two scripts: one where bots can play against each other in a single game where all moves in the game can be seen, and a script where bots play a tournament against each other - with a specifiable number of games to be played - and where the number of won games per robot can be seen. There are already multiple bots present in the framework to play against:
\begin{description}
\item [Minimax / Alphabeta:] This bot only works in a perfect-information game (phase 2 of Schnapsen). It recursively looks through all the states, and decides what the best move to play will be, provided that the opponent plays optimally, and will try to minimize the outcome of the game for the player.
\item [Bully:] Plays by a defined set of rules. If it has a trump card, it will play that. Otherwise, it will play a card of the same suit player 1 played. If it cannot do either of those things, it will play the highest ranking card of any suit.
\item [Rand:] Chooses a random move out of all of the legal moves, and plays that.
\item [Rdeep:] Functions quite like a Monte-Carlo sampling bot. It looks through multiple states, and averages the heuristics of a state, thereby ranking the moves. It then picks one of the highest ranking moves to play.
\item [KBbot:] Logical reasoning bot. The strategy can be defined in propositional logic. The current strategy of the bot is to always play cheap cards (i.e kings, queens, jacks) first.
\end{description}

\subsection{Machine Learning Bots}
For this paper, we will use machine learning bots with different sets of features to determine what the best features are.

@@

To answer our research question, multiple Schnapsen tournaments were held. Each learning bot played multiple 120 game tournaments against each individual non-learning bot.

\subsection{The bots}
The non- learning opponents were Rand, Bully and Rdeep. The learning bots we created were:
\begin{description}
\item [ml\_minimal:] This bot has only the perspective and both players points as features.

\item [ml\_stripped:] ml\_stripped has all the \textit{default feature space} except for both players pending points and trumpsuit id.

\item [ml:] This ml bot is provided by the framework and consists of the  \textit{default feature space}

\item [ml\_enriched:]  This bot has all the features of ml bot. In addition to the \textit{default feature space} the following features were added to the feature space:
\begin{itemize}
\item Both player's points squared
\item Both player's pending points squared
\item Both player's points cubed
\item Both player's pending points cubed
\end{itemize}

\item [ml\_advanced:] ml\_advanced  has the following features in addition to the \textit{default feature space}:
\begin{itemize}
\item Difference between points
\item Difference between pending points
\item Trump card ratio in hand
\item High card ratio in hand (10/A's)
\item Number of trump cards in hand in phase 2
\item Number of cards with same suit as opponents played suit
\item Number of points in hand (squared)
\end{itemize}
\end{description}

\subsection{The testing procedure}

Each learning bot played a total of 9 tournaments with 3 different models against each non-learning bot.
The testing procedure was as follows:
\begin{itemize}
\item Training
\item 120 Game tournament
\item 120 Game tournament
\item 120 Game tournament
\end{itemize}

This procedure was repeated twice to reach the total of 9 tournaments respectively. The testing only occured between learning and non-learning bots.  The learning bots did not compete against eachother.  In case of Rand and Bully, the training process consisted of 100,000 observations of the (non-learning) opponent . Rdeep was only observed 10,000 times during training.
